\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2017-05-17 redaction/dhp_sept2016.tex
changement encore

In this section, we investigate the role of the additive component in total transport costs in a historical perspective. As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and iceberg components in transport costs over time (by transport mode).
Let us consider first the trend in total transport costs (ie, considering the upper line in Figure \ref{fig:decomp_TC_3d}). Both air and ocean shipping exhibit a downward trend in overall transport costs since 1974. (-2.1\% a year for mean air transport costs and -2.0\% a year for mean ocean transport costs).
\footnote{One may worry that this result springs from high oil-shock related transport costs in 1974. However, computing the time trends from 1980 does not change dramatically the picture. The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean ocean transport costs. In order to go as far as possible into the past, this paper will focus on the results from 1974}
 \todo{GD : j'ai amélioré. Qu'en penses-tu ?}
These results are consistent with \citet{hummels2007}.

@@

If we share this view with \citet{hummels2007}, our strategy for eliminating composition effects differs from his in two main respects. First, we provide a broader characterization of the time trends in transport costs, by identifying the trend patterns of the ``pure'' transport costs'' (composition effects excluded) for the two additive and ad-valorem dimensions. Our empirical analysis indeed allows for a flexible estimation of the share between the additive and the ad-valorem components of transport costs, over time/product/country partner.\footnote{The empirical specification of \cite{hummels2007}, as in \cite{hummels_skiba}, assumes a constant share of additive transport costs over time, product and country partner. In \cite{hummels_skiba}, this shows up through the fact that the elasticity of freight costs to prices $\beta$ is estimated as a constant (with $0<\beta<1$, the lower $\beta$ the larger the additive cost dimension). By contrast, our estimation strategy allows for a varying $\beta$ across time, country partner and product.} Accordingly, our estimation strategy allows for an explicit decomposition of the additive cost component taking these three dimensions into account. By contrast, in its decomposition exercise to characterize the time trend in transport costs, \citet{hummels2007} does not take into account the changes in their additive dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \citet{hummels2007}, p. 146. } By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time. We proceed similarly for the multiplicative component of trade costs and, for sake of completeness, on the observed cif-fas price gap.

>>>>

In this section, we investigate the role of the additive component in total transport costs in a historical perspective. As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and iceberg components in transport costs over time (by transport mode).
Let us consider first the trend in total transport costs (ie, considering the upper line in Figure \ref{fig:decomp_TC_3d}). Two main comments can be made. First, both air and ocean shipping exhibit a downward trend in overall transport costs since 1974 (-2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs).\footnote{One may worry that this result springs from high oil-shock related transport costs in 1974. However, computing the time trends from 1980 does not dramatically change the picture. The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean ocean transport costs. We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.} Second, the decrease appears more pronounced for sea than air transportation. These results are consistent with \citet{hummels2007}.

@@

If we share this view with \citet{hummels2007}, our strategy for eliminating composition effects differs from his in two main respects. First, we provide a broader characterization of the time trends in transport costs, by identifying the trend patterns of the ``pure'' transport costs'' (composition effects excluded) for the two additive and ad-valorem dimensions. Our empirical analysis indeed allows for a flexible estimation of the share between the additive and the ad-valorem components of transport costs, over time/product/country partner.\footnote{The empirical specification of \cite{hummels2007}, as in \cite{hummels_skiba}, assumes a constant share of additive transport costs over time, product and country partner. In \cite{hummels_skiba}, this shows up through the fact that the elasticity of freight costs to prices $\beta$ is estimated as a constant (with $0<\beta<1$, the lower $\beta$ the larger the additive cost dimension). By contrast, our estimation strategy allows for a varying $\beta$ across time, country partner and product.} Accordingly, our estimation strategy allows for an explicit decomposition of the additive cost component taking these three dimensions into account. By contrast, in its decomposition exercise to characterize the time trend in transport costs, \citet{hummels2007} does not take into account the changes in their additive dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \citet{hummels2007}, p. 146. } By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time. We proceed similarly for the multiplicative component of trade costs. We also agglomerate the two components in a unified measure of transport costs from which we exclude the composition effects. This
allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones.\footnote{Notice that these ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported in another perspective (basis 100 in year 1974).}
